---
title: "Using BayesTransitionModels to generate national estimates of mCPR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using BayesTransitionModels to generate national estimates of mCPR}
  %\VignetteEngine{knitr::knitr}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6.5,
  fig.height = 5
)
```

This vignette demonstrates fitting the `BayesTransitionModels` model to national modern contraceptive prevalence rate (mCPR) data from a subset of countries.

```{r setup}
library(tidyverse)
library(BayesTransitionModels)
```

To start, we use the `national_data` function to extract national mCPR data from 10 countries.
```{r extract_data}
data <- national_data(
  countries = c("Somalia", "India", "Palau", "Bangladesh", "Zimbabwe",
                "Indonesia", "Kenya", "Turkey", "Mexico", "Guatemala"),
  start_year = 1970
)
```

Let's take a look at what the raw data look like.
```{r data_plot}
data %>%
  ggplot(aes(x = year, y = contraceptive_use_modern)) +
  geom_point() +
  facet_wrap(~name_country) +
  labs(x = "Year", y = "mCPR")
```

Next, we fit the model using the `BayesTransitionModels` function. Note here for demonstration purposes, and to save computation time, we are using values for `adapt_delta` and `max_treedepth` that are too low and cause divergent transitions and transitions that exceed the maximum treedepth; in practice both of these parameters should be increased until no warning messages are shown.

```{r bayestransitionmodels-fit, cache = TRUE}
fit <- transitionmodel(
  data,
  y = "contraceptive_use_modern",
  se = "se_modern",
  year = "year",
  source = "data_series_type",
  area = "name_country",
  
  start_year = 1970, 
  end_year = 2030, 
  
  model = "spline",
  
  # Spline setup
  spline_degree = 2,
  num_knots = 7,
  
  # Hierarchical setup
  hierarchical_level     = c("intercept", "name_country"), 
  hierarchical_splines   = c("intercept", "name_country"),
  hierarchical_asymptote = c("intercept", "name_country"),
  
  # Prior settings
  tau_prior = "normal(0, 2)",
  rho_prior = "uniform(0, 1)",
  
  # Stan sampler settings
  adapt_delta = 0.95,
  max_treedepth = 8,
  iter_warmup = 200,
  iter_sampling = 200,
  seed = 5,
  parallel_chains = 4,
  refresh = 50
)
``` 

Calling `summary` displays some basic information about the model fit.
```{r transitionmodel_summary}
summary(fit)
```

Now let's plot the main results. Use `plot_indicator` to plot the estimates and projections of mCPR over time in each country. The black line shows the posterior median of the true mCPR in each country ($\eta_{c,t}$) and time point and the shaded regions show the 50%, 80%, and 95% posterior credible intervals.

```{r plot_indicator}
plot_indicator(fit)
```

The `plot_smoother` function can be used to view the posterior estimates of the smoothing component ($\eta_{c,t}$).
```{r plot_smoother}
plot_smoother(fit)
```

The transition function can be plotted for each individual country using `plot_transition`. We can also plot the hierarchical global mean transition function by specifying `"intercept"` as the hierarchical level.
```{r plot_transitions}
plot_transition(fit, hierarchical_level = "name_country")
plot_transition(fit, hierarchical_level = "intercept")
```

Next, the asymptotes ($P_c$) can be plotted for each level of hierarchy.
```{r plot_asymptote_country}
plot_asymptote(fit, "name_country", sort = TRUE)
```

```{r plot_asymptote_intercept, fig.height = 2.5}
plot_asymptote(fit, "intercept")
```

Similarly, we can plot the level in reference year ($\Omega_c$) for each country and for the hierarchical global mean.
```{r plot_level_country}
plot_level(fit, hierarchical_level = "name_country", sort = TRUE)
```

```{r plot_level_intercept, fig.height = 2.5}
plot_level(fit, hierarchical_level = "intercept")
```

The hyperparameters of the smoothing model ($\rho$ and $\tau$) can also be inspected.
```{r plot_smoothing_hyperparameters, fig.height = 4}
plot_smoothing_hyperparameters(fit, priors = list(
  tau = function(x) truncnorm::dtruncnorm(x, a = 0, b = Inf, mean = 0, sd = 2),
  rho = function(x) dunif(x, 0, 1)
))
```

```{r plot_data_hyperparameters}
plot_data_hyperparameters(fit, include_prior = TRUE)
```
